---
title: "Tidytuesday"
author: "Kevin Wang"
date: "`r paste0('Initiated on 2019 May 21, compiled on ', format(Sys.time(), '%Y %b %d'))`"
output:
  html_document:
    code_folding: hide
    fig_height: 12
    fig_width: 12
    toc: yes
    number_sections: true
    toc_depth: 3
    toc_float: yes
editor_options: 
  chunk_output_type: console
---

# Summary 

+ This data originates from the `tidytuesday` challenge. (https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-05-21)

+ There are three pieces of data, however, I will only focus on one of these: `waste_vs_gdp`. 

+ I speculate the amount of waste is correlated to the GDP but also **GDP growth** (https://twitter.com/search?q=tidytuesday%20plastic&src=typd). Since GDP growth is not provided, I have decided to download the data from World Bank. 





# Loading packages
```{r}
library(tidyverse)
library(janitor)
library(gganimate)
library(skimr)
library(ggsci)

theme_set(theme_classic(18) + 
            theme(legend.position = "bottom"))
```


# Loading data
```{r}
coast_vs_waste <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-05-21/coastal-population-vs-mismanaged-plastic.csv") %>% 
  janitor::clean_names()

mismanaged_vs_gdp <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-05-21/per-capita-mismanaged-plastic-waste-vs-gdp-per-capita.csv") %>% 
  janitor::clean_names()

waste_vs_gdp <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-05-21/per-capita-plastic-waste-vs-gdp-per-capita.csv") %>% 
  janitor::clean_names()
```

# Cleaning up `waste_vs_gdp`
```{r}
waste_vs_gdp_missing = waste_vs_gdp %>% 
  group_by(entity) %>% 
  summarise(nMissing = per_capita_plastic_waste_kilograms_per_person_per_day %>% is.na %>% sum) %>% 
  dplyr::arrange(desc(nMissing))



waste_vs_gdp_cleaned = waste_vs_gdp %>% 
  dplyr::transmute(
    entity, code, year, 
    plastic_standard = per_capita_plastic_waste_kilograms_per_person_per_day,
    pop = total_population_gapminder) %>% 
  na.omit()

waste_vs_gdp_cleaned
```


# Loading World Bank data 

```{r}
gdp_growth_raw = readr::read_csv("API_NY.GDP.PCAP.KD.ZG_DS2_en_csv_v2_10576663/API_NY.GDP.PCAP.KD.ZG_DS2_en_csv_v2_10576663.csv", skip = 4)

country_regions = readr::read_csv("API_NY.GDP.PCAP.KD.ZG_DS2_en_csv_v2_10576663/Metadata_Country_API_NY.GDP.PCAP.KD.ZG_DS2_en_csv_v2_10576663.csv", skip = 0) %>% 
  dplyr::select(entity = TableName, 
                region = Region)



gdp_growth_cleaned = gdp_growth_raw %>% 
  dplyr::rename(entity = `Country Name`
                # code = `Country Code`
) %>% 
  dplyr::select(entity, `1960`:`2018`) %>% 
  tidyr::gather(key = growth_year,
                value = gdp_growth, 
                `1960`:`2018`) %>% 
  left_join(country_regions, by = "entity")

gdp_growth_cleaned %>% skimr::skim()


gdp_growth_nest = gdp_growth_cleaned %>% 
  group_by(entity) %>% 
  tidyr::nest()


waste_join_growth = waste_vs_gdp_cleaned %>% 
  inner_join(gdp_growth_nest, by = "entity") %>% 
  unnest() %>% 
  dplyr::filter(growth_year >= 2001, growth_year != 2018)

waste_join_growth %>% skimr::skim()

waste_join_growth_filled = waste_join_growth %>% 
  group_by(entity) %>% 
  dplyr::arrange(growth_year) %>% 
  tidyr::fill(gdp_growth, .direction = "up") %>% 
  dplyr::filter(abs(gdp_growth) <= 20, plastic_standard < 2, complete.cases(region))
# dplyr::mutate(growth_year = growth_year %>% ISOdate(1, 1) %>% as.Date)
```


# GDP growth vs waste in 2010 only

```{r}
waste_join_growth_filled_2010 = waste_join_growth_filled %>% 
  dplyr::filter(growth_year == 2010, abs(gdp_growth) <= 20, plastic_standard < 2)



waste_join_growth_filled_2010 %>% 
  ggplot(aes(x = gdp_growth, y = plastic_standard)) +
  geom_point(aes(colour = region), size = 3) +
  geom_smooth(aes(group = 1), method = "lm", se = FALSE) +
  ggsci::scale_color_d3()
```



# GDP growth vs waste over (extropolated years)

```{r, message=FALSE}
waste_join_growth_filled_lm = waste_join_growth_filled %>% 
  group_by(growth_year) %>% 
  tidyr::nest() %>% 
  dplyr::mutate(
    lm_tidy = purrr::map(
      .x = data, 
      .f = ~ lm(plastic_standard ~ gdp_growth, data = .x) %>% 
        broom::tidy()),
    slope_est = purrr::map_dbl(
      lm_tidy, .f = ~ .x$estimate[2]) %>% 
      round(4) %>% paste0("Slope = ", .), 
    slope_pval = purrr::map_dbl(
      lm_tidy, .f = ~ .x$p.value[2]) %>% 
      round(4) %>% paste0("Slope pvalue = ", .), 
    corr_est = purrr::map_dbl(
      .x = data, 
      .f = ~ cor(.x$plastic_standard, .x$gdp_growth)) %>% 
      round(4) %>% paste0("Corr = ", .), 
  ) %>%
  dplyr::select(-lm_tidy)
# tidyr::unnest()

waste_join_growth_filled_lm





gif = waste_join_growth_filled %>%
  ggplot(aes(x = gdp_growth,
             y = plastic_standard)) +
  geom_point(aes(colour = region, group = 1L), size = 3) +
  geom_smooth(aes(group = growth_year), method = "lm", se = FALSE) +
  geom_text(data = waste_join_growth_filled_lm,
            mapping = aes(x = -10, y = 0.6, label = slope_est), 
            colour = "red", size = 5) +
  geom_text(data = waste_join_growth_filled_lm,
            mapping = aes(x = -10, y = 0.5, label = slope_pval), 
            colour = "red", size = 5) +
  geom_text(data = waste_join_growth_filled_lm,
            mapping = aes(x = -10, y = 0.4, label = corr_est), 
            colour = "red", size = 5) +
  ggsci::scale_colour_d3() +
  transition_states(growth_year,
                    transition_length = 0.1,
                    state_length = 0.5) +
  labs(title = 'Year: {closest_state}',
       x = 'GDP growth per capita',
       y = 'Plastic waste per capita (kg/day), in 2010') +
  # theme(title = slope_est) +
  # shadow_mark() + ## Make points stay
  ease_aes('linear')
# facet_wrap(~growth_year)

anim_save(filename = "gdp_growth_vs_plastic_per_cap.gif",
          animation = gif, width = 800)
```





# Session Info
```{r}
sessionInfo()
```

