---
title: "Animal rescues"
author: "Kevin Wang"
date: "`r paste0('Initiated on 2021 June 29, compiled on ', format(Sys.time(), '%Y %b %d'))`"
output:
  html_document:
    code_folding: hide
    fig_height: 6
    fig_width: 10
    toc: yes
    number_sections: true
    toc_depth: 3
    toc_float: yes
    theme: paper
editor_options: 
  chunk_output_type: console
---

```{r}
knitr::opts_chunk$set(warning = FALSE,
                      message = FALSE)
```


```{r}
library(tidyverse)
library(lubridate)

theme_set(theme_classic(18) +
            theme(legend.position = "bottom"))

# download.file(url = 'https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-06-29/animal_rescues.csv', destfile = "./animal_rescues.csv")
animal_rescues <- readr::read_csv(file = "./animal_rescues.csv")
```

```{r}
skimr::skim(animal_rescues)
```

# Original animal grouping
```{r}
table(animal_rescues$animal_group_parent) %>% knitr::kable()
```

# Selecting columns/cleaning up data for visualisation

The main improvements are: 

+ Coerced date time.
+ Animal are collapsed into 8 types only for easy visualisations.
+ Error entry of a bird with latitude and longitude of zero was removed.

```{r}
subdata = animal_rescues %>% 
  dplyr::transmute(
    incident_number, 
    date_time_of_call = lubridate::dmy_hm(date_time_of_call),
    animal_type = animal_group_parent %>% tolower,
    animal_type = case_when(
      str_detect(animal_type, "farm animal") ~ "farm animal", 
      str_detect(animal_type, "heavy livestock animal") ~ "farm animal", 
      animal_type == "bull" ~ "cow",
      animal_type == "budgie" ~ "bird",
      animal_type == "pigeon" ~ "bird",
      # str_detect(animal_type, "unknown") ~ "unknown",
      TRUE ~ animal_type),
    animal_type = fct_lump(f = animal_type, n = 8),
    animal_type_int = animal_type %>% as.factor() %>% as.integer(),
    latitude = as.numeric(latitude), 
    longitude = as.numeric(longitude)) %>% 
  dplyr::filter(complete.cases(latitude), complete.cases(longitude), latitude > 0)

# subdata
# glimpse(subdata)
table(subdata$animal_type) %>% knitr::kable()
```

# Mapbox visualisation using `plotly`

Click legend to select for specific type of animals.

```{r}
library(plotly)
Sys.setenv('MAPBOX_TOKEN' = Sys.getenv('MAPBOX_AUTH'))

fig = subdata %>% 
  plot_mapbox(lat = ~latitude, lon = ~longitude,
              split = ~animal_type, 
              size = 2,
              mode = 'scattermapbox', 
              hoverinfo = 'animal_type')

fig = fig %>% 
  layout(title = 'Rescue animal types',
         font = list(color='white'),
         plot_bgcolor = '#191A1A', 
         paper_bgcolor = '#191A1A',
         mapbox = list(style = 'dark', 
                       zoom = 7,
                       center = list(
                         lat = median(subdata$latitude),
                         lon = median(subdata$longitude))),
         legend = list(orientation = 'h',
                       font = list(size = 8)),
         margin = list(l = 25, r = 25,
                       b = 25, t = 25,
                       pad = 2)) 

fig
```

# Saving `subdata` for visulisation in the main page 
```{r, eval = FALSE}
saveRDS(subdata, file = "sub_animal_rescues.rds")
```

<!-- # Time visualisation -->
<!-- ```{r} -->
<!-- subdata %>%  -->
<!--   # dplyr::filter(animal_type == "fox") %>%  -->
<!--   dplyr::mutate(hour = lubridate::hour(date_time_of_call), -->
<!--                 quater = lubridate::quarter(date_time_of_call)) %>%  -->
<!--   ggplot(aes(x = hour)) + -->
<!--   # geom_histogram() + -->
<!--   geom_density(aes(colour = factor(quater))) + -->
<!--   facet_wrap(~animal_type) -->
<!-- ``` -->


<!-- # Mapbox visualisation using `mapdeck` -->

<!-- ```{r, eval = FALSE} -->
<!-- library(mapdeck) -->
<!-- set_token(Sys.getenv("MAPBOX_AUTH")) -->
<!-- # mapdeck(token = mapdeck(token = 'your_token')) -->
<!-- mapdeck::mapdeck(style = mapdeck_style('dark')) %>%  -->
<!--   mapdeck::add_scatterplot( -->
<!--     data = subdata, -->
<!--     lat = "latitude", -->
<!--     lon = "longitude", -->
<!--     tooltip = "animal_type", -->
<!--     radius = 200, -->
<!--     fill_colour = "animal_type", -->
<!--     palette = "viridis", -->
<!--     legend = TRUE -->
<!--   ) -->
<!-- ``` -->

