---
title: "Bird Impacts"
author: "Kevin Wang"
date: "5/30/2019"
output:
  html_document:
    code_folding: hide
    fig_height: 10
    fig_width: 10
    toc: yes
    number_sections: true
    toc_depth: 3
    toc_float: yes
editor_options: 
  chunk_output_type: console
---

# Loading the packages
```{r}
library(tidyverse)
library(janitor)
library(skimr)
library(maps)
library(gganimate)
library(ggpubr)

theme_set(theme_classic(18) +
            theme(legend.position = "bottom"))
```



# Load data

```{r}
raw_data = read_csv("bird_impacts.csv") %>% 
  janitor::clean_names() %>% 
  dplyr::mutate(incident_month = factor(incident_month), 
                incident_year = factor(incident_year), 
                month = lubridate::month(incident_date, label = TRUE, abbr = TRUE))

raw_data %>% 
  skimr::skim()
```


This data has a large number of missing data in some fields. It seems like we should consider analysing individual variables 

# Does particular months has more birds impacts?

We put forward the hypothesis that some months has more impacts due to weather and bird behaviours. 

We will also visualise the number of incidence per year per month to detect any report-time bias in the data. 


```{r}
month_summ = raw_data %>% 
  dplyr::group_by(incident_year, month) %>% 
  dplyr::count()


month_trace_plot = ggplot(month_summ, 
                          aes(x = month, 
                              y = n, 
                              colour = incident_year, 
                              group = incident_year)) +
  geom_line() +
  labs(title = "Number of birds impacts per year per month",
       y = "Number of incidents") +
  theme(legend.title = element_blank()) +
  guides(colour = guide_legend(nrow = 3, byrow = TRUE))

month_trace_plot
```

These numbers should be scaled by the number of flights in the US every year. This analysis will be delayed. 

It should be clear that the number of impacts are lower in the winter (Dec to Feb) and both spring and autumn has a higher number of impacts. It is interesting that there is a dip in June every year. This could be caused by some sampling bias or somthing about bird behaviours. 


# Geographical analysis of incidence in 2018
```{r}
state_data = tibble::tibble(
  state_abb = datasets::state.abb, 
  state_name = tolower(datasets::state.name),
  state_division = datasets::state.division
) 
#########################

state_summ = raw_data %>% 
  dplyr::filter(incident_year == 2018) %>% 
  dplyr::mutate(state = factor(state)) %>% 
  dplyr::group_by(month, state, .drop = FALSE) %>% 
  dplyr::tally() %>% 
  ungroup() %>% 
  left_join(state_data, by = c("state" = "state_abb"))

state_summ

summary(state_summ$n)

#################
us_states <- map_data("state") %>% 
  left_join(state_summ, by = c("region" = "state_name")) %>% 
  dplyr::filter(complete.cases(month))

ggplot(data = us_states,
       mapping = aes(x = long, y = lat,
                     group = group)) +
  geom_polygon(aes(fill = n), colour = "black") +
  facet_wrap(~month) +
  scale_fill_distiller(palette = "Reds", direction = 1) +
  labs(title = "No. of bird impacts in 2018") +
  theme_void(18) +
  theme(legend.position = "bottom")
```

https://socviz.co/maps.html


## gganimate version
```{r}
gif = ggplot(data = us_states,
             mapping = aes(x = long, y = lat,
                           group = group)) +
  geom_polygon(aes(fill = n), colour = "black") +
  scale_fill_distiller(palette = "Reds", direction = 1) +
  theme_void(18) +
  theme(legend.position = "bottom") +
  transition_states(month,
                    transition_length = 0.1,
                    state_length = 0.5) +
  ease_aes('linear') +
  labs(title = "No. of bird impacts in 2018",
       subtitle = "Month: {closest_state}")


anim_save(filename = "incidents_2018.gif",
          animation = gif, width = 800)

```

# Are there more impacts at night? 
```{r}
time_summ = raw_data %>% 
  dplyr::filter(complete.cases(time_of_day)) %>% 
  dplyr::group_by(incident_year, month, time_of_day) %>% 
  dplyr::count()


time_trace_plot = ggplot(time_summ, 
                         aes(x = month, 
                             y = n, 
                             colour = incident_year, 
                             group = interaction(incident_year, time_of_day))) +
  geom_line() +
  facet_grid(~ time_of_day) +
  labs(title = "Number of birds incidents during the time of day",
       y = "Number of incidents") +
  theme(legend.title = element_blank(), 
        axis.text.x = element_text(angle = 90),
        legend.position = "none")

time_trace_plot
```


Interestingly, the dipping effect in June was only visible at night times. 


```{r}
combine_trace_plot = cowplot::plot_grid(
  month_trace_plot,
  time_trace_plot,
  nrow = 2,
  rel_heights = c(2, 1))

combine_trace_plot

ggsave(plot = combine_trace_plot, 
       filename = "combine_trace_plot.png",
       width = 12, height = 8)
```



# Height and incidence rate

```{r}
complete_phase = raw_data %>% 
  dplyr::filter(complete.cases(phase_of_flt)) %>% 
  dplyr::mutate(
    phase_of_flt = tolower(phase_of_flt),
     phase_of_flt_cat = phase_of_flt %>% 
      forcats::fct_lump(n = 15) %>% 
      forcats::fct_lump_min(min = 5) %>% 
       forcats::fct_infreq(),
     log_height = log10(height + 1)
  )

table(complete_phase$phase_of_flt_cat)

complete_phase %>% 
  ggplot(aes(x = phase_of_flt_cat)) +
  geom_bar()


complete_phase %>% 
  ggplot(aes(x = log_height)) +
  geom_histogram() +
  facet_wrap(~phase_of_flt_cat, scales = "free")

complete_phase %>% 
  dplyr::filter(log_height > 0) %>% 
  ggplot(aes(x = log_height)) +
  geom_histogram(bins = 30) +
  facet_wrap(~phase_of_flt_cat, scales = "free")
```



## Visualisation on cost of repair
```{r}
table(is.na(complete_phase$cost_repairs_infl_adj))
summary(complete_phase$cost_repairs_infl_adj)
hist(complete_phase$cost_repairs_infl_adj)

complete_cost_data = complete_phase %>% 
  dplyr::filter(complete.cases(cost_repairs_infl_adj)) %>% 
  dplyr::rename(cost_repairs = cost_repairs_infl_adj) %>% 
  dplyr::mutate(
    atype_cat = atype %>% 
      forcats::fct_lump(n = 15) %>% 
      forcats::fct_lump_min(min = 5) %>%
      forcats::fct_reorder(cost_repairs, .fun = median),
    species_cat = species %>% 
      forcats::fct_lump(n = 20) %>% 
      forcats::fct_lump_min(min = 5) %>%
      forcats::fct_infreq(),
    phase_of_flt = forcats::fct_infreq(phase_of_flt)
    ) %>% 
  tidyr::separate(atype,
                  into = c("acompany", "amodel", "avariant"), 
                  sep = "-", 
                  remove = FALSE)

complete_cost_data %>% 
  group_by(atype_cat) %>% 
  summarise(median(cost_repairs))


complete_cost_data %>% 
  dplyr::filter(atype_cat != "Other") %>% 
  ggplot(aes(x = atype_cat, 
             y = cost_repairs,
             fill = acompany)) +
  # geom_violin(width = 1) +
  geom_boxplot(width = 0.5) +
  scale_y_log10(label = scales::comma) +
  ggsci::scale_fill_nejm() +
  coord_flip()


complete_cost_data %>% 
  dplyr::filter(species_cat != "Other") %>% 
  ggplot(aes(x = species_cat, 
             y = cost_repairs)) +
  # geom_violin(width = 1) +
  geom_boxplot(width = 0.5) +
  scale_y_log10(label = scales::comma) +
  ggsci::scale_fill_nejm() +
  coord_flip()


complete_cost_data %>% 
  group_by(phase_of_flt, species_cat, .drop = FALSE) %>% 
  dplyr::count() %>% 
  ggplot(aes(x = phase_of_flt, 
             y = species_cat, 
             fill = n,
             label = n)) +
  geom_tile() +
  geom_text() +
  scale_fill_distiller(palette = "Spectral")


complete_cost_data %>% 
  ggplot(aes(x = month, 
             y = cost_repairs)) +
  # geom_violin(width = 1) +
  geom_boxplot(width = 0.5) +
  scale_y_log10(breaks = c(1e4, 2e4, 5e4, 7e4, 
                           1e5, 2e5, 5e5, 7e5, 
                           1e6, 2e6, 5e6, 7e6), 
                label = scales::comma)
```



```{r}
library(ggparty)

complete_cost_data %>% skimr::skim()

cost_data_reg = complete_cost_data %>% 
  dplyr::filter(complete.cases(sky), 
                # complete.cases(phase_of_flt),
                complete.cases(speed), 
                complete.cases(height),
                complete.cases(damage),
                complete.cases(time_of_day)) %>% 
  dplyr::transmute(
    sky, 
    # phase_of_flt, 
    damage_three = forcats::fct_recode(damage, "M" = "M?") %>% 
      forcats::fct_relevel("N"),
    speed, 
    log_height = log10(height + 1L), 
    time_of_day, 
    log_cost_repairs = log10(cost_repairs),
  ) %>% as.data.frame()


lm(log_cost_repairs ~ ., data = cost_data_reg) %>% 
  broom::tidy()

library(randomForest)

rf = randomForest::randomForest(
  x = model.matrix(log_cost_repairs ~., data = cost_data_reg),
  y = cost_data_reg$log_cost_repairs)

varImpPlot(rf)


cost_data_reg %>% 
  ggplot(aes(x = log_height, 
             y = log_cost_repairs,
             colour = damage_three)) +
  geom_point() +
  geom_smooth(method = "loess")

cost_data_reg %>% 
  dplyr::filter(log_height > 0) %>% 
  ggplot(aes(x = log_height, 
             y = log_cost_repairs,
             colour = damage_three)) +
  geom_point() +
  geom_smooth(method = "loess")
```



```{r}
library(ggparty)

cost_data_reg = cost_data_reg %>%
  dplyr::mutate_if(is.character, as.factor)

tree <- ctree(log_cost_repairs ~ ., data = cost_data_reg)

ggparty(tree, horizontal = TRUE, terminal_space = 0.6) +
  geom_edge() +
  geom_edge_label() +
  geom_node_splitvar() +
  geom_node_plot(gglist = list(
    geom_density(aes(x = log_cost_repairs))),
    shared_axis_labels = TRUE)
```



https://cran.r-project.org/web/packages/geofacet/index.html
# Session Info

```{r}
sessionInfo()
```

