---
title: "Collinearity in Test cricketer data"
author: "Kevin Wang"
date: "16/08/2019"
output:
  html_document:
    code_folding: hide
    fig_height: 10
    fig_width: 10
    toc: yes
    number_sections: true
    toc_depth: 3
    toc_float: yes
editor_options: 
  chunk_output_type: console
---

# Introduction

In this document, we will visualise the cleaned cricket

```{r}
library(tidyverse)
library(mcvis)
```


# Loading data
```{r}
clean_test_batting_career = readr::read_csv("./clean_test_batting_career.csv")

clean_test_batting_inngings = readr::read_csv("./clean_test_batting_inngings.csv")
```


# Confirm relationships

```{r}
plot(clean_test_batting_career$strike_rate/100,
     clean_test_batting_career$runs/clean_test_batting_career$balls_faced)
abline(a = 0, b = 1, col = "red")

plot(clean_test_batting_career$ave,
     clean_test_batting_career$runs/(clean_test_batting_career$inns - clean_test_batting_career$not_out))
abline(a = 0, b = 1, col = "red")
```


# Correlation betwen key numeric variables

Looking at the correlation plot, it is not clear as to which variable is the main culprit that causes multi-collinearity. 

## Corrplot
```{r}
X = clean_test_batting_career %>% 
  dplyr::mutate(hs_num = hs %>% stringr::str_extract("\\-*\\d+\\.*\\d*") %>% as.integer()) %>% 
  dplyr::select_if(is.numeric) %>% 
  na.omit() %>% 
  dplyr::transmute(
    # log_ave = log10(ave),
    log_sr = log10(strike_rate),
    log_bf = log10(balls_faced/100),
    # log_outs = log10(inns - not_out),
    log_fours = log10(fours + 1L),
    log_ducks = log10(ducks + 1L),
    log_sixes = log10(sixes + 1L),
    log_hs = log10(hs_num),
    log_hcentury = log10(half_century + 1L),
    log_runs = log10(runs)
    # log_inns = log10(inns + 1L),
    # log_not_out = log10(not_out + 1L)
    ) 



skimr::skim(X)

plot(X$log_sr, 
     X$log_runs - X$log_bf)
abline(a = 0, b = 1, col = "red")


# plot(X$log_ave, 
#      X$log_sr + X$log_bf - X$log_outs)
# abline(a = 0, b = 1, col = "red")



skimr::skim(X)

corrplot::corrplot.mixed(
  cor(X), 
  upper = "square")
```

## d3heatmap
```{r}
d3heatmap::d3heatmap(cor(X))
```

## Scatter plot
```{r, fig.height=12,fig.width=12}
X %>% 
  pairs(lower.panel = CPOP::panel_cor)
```


# Linear regression
```{r}
m = lm(log_sr ~ ., data = X)

summary(m)
```


# VIF

```{r}
library(car)
m %>% 
  car::vif() %>% 
  round(2)
```



# mcvis


mcvis method seems to point to the `runs` variable as the main cause of colinearity.

```{r}
mcvis_result = X %>% 
  # dplyr::select(-log_sr) %>% 
  mcvis::mcvis(standardise_method = "none")

mcvis_result$MC %>% round(2)

mcvis::ggplot_mcvis(mcvis_result)
```


# Session Info
```{r}
sessionInfo()
```

