---
title: "Collinearity in Test cricketer data"
author: "Kevin Wang"
date: "16/08/2019"
output:
  html_document:
    code_folding: hide
    fig_height: 10
    fig_width: 10
    toc: yes
    number_sections: true
    toc_depth: 3
    toc_float: yes
editor_options: 
  chunk_output_type: console
---

# Introduction

In this document, we will visualise the cleaned cricket

```{r}
library(tidyverse)
library(mcvis)
```


# Loading data
```{r}
clean_test_batting_career = readr::read_csv("./clean_test_batting_career.csv")

clean_test_batting_inngings = readr::read_csv("./clean_test_batting_inngings.csv")
```


# Confirm relationships

```{r}
plot(clean_test_batting_career$strike_rate/100,
     clean_test_batting_career$runs/clean_test_batting_career$balls_faced)
abline(a = 0, b = 1, col = "red")

plot(clean_test_batting_career$ave,
     clean_test_batting_career$runs/(clean_test_batting_career$inns - clean_test_batting_career$not_out))
abline(a = 0, b = 1, col = "red")
```


# Correlation betwen key numeric variables

Looking at the correlation plot, it is not clear as to which variable is the main culprit that causes multi-collinearity. 

## Corrplot
```{r}
X = clean_test_batting_career %>% 
  dplyr::select_if(is.numeric) %>% 
  dplyr::select(-fours, -half_century, -century) %>% 
  na.omit() %>% 
  dplyr::mutate(
    log_ave = log10(ave),
    log_sr = log10(strike_rate),
    log_bf = log10(balls_faced/100),
    log_denom = log10(inns - not_out)
    # log_runs = log10(runs + 1L),
    # log_inns = log10(inns + 1L),
    # log_not_out = log10(not_out + 1L)
    ) %>% 
dplyr::select(-career_start,
              -career_end,
              -ave,
              -strike_rate,
              -balls_faced,
              -inns, 
              -not_out,
              -mat,
              -runs)

glimpse(X)

plot(X$log_ave, 
     X$log_sr + X$log_bf - X$log_denom)
abline(a = 0, b = 1, col = "red")



skimr::skim(X)

corrplot::corrplot.mixed(
  cor(X), 
  upper = "square")
```

## d3heatmap
```{r}
d3heatmap::d3heatmap(cor(X))
```


# mcvis


mcvis method seems to point to the `runs` variable as the main cause of colinearity.

```{r}
mcvis_result = X %>% 
  mcvis::mcvis()

mcvis_result$MC %>% round(2)

mcvis::ggplot_mcvis(mcvis_result)
```


# Session Info
```{r}
sessionInfo()
```

