---
title: "Threholding analysis of cricketer scores"
author: "Kevin Wang"
date: "27/08/2019"
output:
  html_document:
    code_folding: hide
    fig_height: 10
    fig_width: 10
    toc: yes
    number_sections: true
    toc_depth: 3
    toc_float: yes
editor_options: 
  chunk_output_type: console
---

# Introduction

In this document, we will try to understand the thresholded cricketer scores. 

The hypothesis is that, after a cricketer scores above a certain number, he will likely score more (after the initial score is subtracted). 

```{r}
library(tidyverse)
```


# Loading data
```{r}
clean_test_batting_inngings = readr::read_csv("./clean_test_batting_inngings.csv")
```


# Data cleaning 

```{r}
inns_data = clean_test_batting_inngings %>% 
  dplyr::mutate(
    not_out = runs %>% stringr::str_detect("[*]") | runs %>% stringr::str_detect("DNB"),
    out = !not_out,
    runs_num = runs %>% stringr::str_extract("\\-*\\d+\\.*\\d*") %>% as.integer() %>% 
      coalesce(0L)
  )

# thres_data %>% 
#   group_by(player_name) %>% 
#   summarise(career_avg = sum(runs_num)/sum(out))
```


## Threshold at 30 check
```{r}
thres_data_30 = inns_data %>% 
  dplyr::mutate(
    runs_num_30 = ifelse(runs_num - 30L < 0,
                         0, runs_num - 30L)) %>% 
  group_by(player_name) %>%
  summarise(career_avg = sum(runs_num_30)/sum(out))
```



# Only keeping the best 100 batsmen in cricket avg
```{r}
career_data = inns_data %>%
  group_by(player_name) %>%
  dplyr::summarise(
    career_avg = sum(runs_num)/sum(out),
    inns = n()) 

career_data_top100 = career_data %>% 
  dplyr::filter(inns >= 20L) %>% 
  dplyr::top_n(n = 20, wt = career_avg)
```


## Threholding at different number of runs
```{r}
inns_data_top100 = inns_data %>% 
  dplyr::filter(player_name %in% career_data_top100$player_name)


thres_career_avg = function(data, thres){
  thres_data = data %>% 
  dplyr::mutate(
    runs_num_thres = ifelse(runs_num - thres < 0,
                         0, runs_num - thres)) %>% 
  group_by(player_name) %>%
  summarise(career_avg_thres = sum(runs_num_thres)/sum(out))
  
  return(thres_data)
}

# thres_career_avg(data = inns_data_top100, thres = 30L)

thres = 0:30
list_thres_career_avg = purrr::map(
  .x = thres, 
  .f = ~ thres_career_avg(data = inns_data_top100, thres = .x)
)
```

## Visualisation
```{r}
mat_thres_career_avg = list_thres_career_avg %>% 
  purrr::map("career_avg_thres") %>% 
  do.call(cbind, .)

df_thres_career_avg = data.frame(mat_thres_career_avg) %>% 
  tibble::as_tibble()
colnames(df_thres_career_avg) = sprintf("%02d", thres)

df_thres_career_avg = df_thres_career_avg %>% 
  dplyr::mutate(player_name = list_thres_career_avg[[1]]$player_name)

df_thres_career_avg_long = df_thres_career_avg %>% 
  tidyr::gather(key = thres, 
                value = career_avg_thres, 
                -player_name) %>% 
  group_by(thres) %>% 
  dplyr::mutate(career_avg_thres_rank = rank(career_avg_thres))


df_thres_career_avg_long %>% 
  dplyr::filter(player_name != "DG Bradman") %>% 
  ggplot(aes(x = thres, 
             y = career_avg_thres,
             colour = player_name)) +
  geom_point() +
  geom_line(aes(group = player_name)) +
  theme(legend.position = "none")
  # scale_colour_brewer(palette = "Set1")

df_thres_career_avg_long %>% 
  # dplyr::filter(player_name != "DG Bradman") %>% 
  ggplot(aes(x = thres, 
             y = career_avg_thres_rank,
             colour = player_name)) +
  geom_point() +
  geom_line(aes(group = player_name)) +
  theme(legend.position = "none")
  # scale_colour_brewer(palette = "Set1")

plotly::ggplotly(last_plot())
```


# Session Info
```{r}
sessionInfo()
```

