---
title: Am I less happy during my PhD?
author: Kevin Wang
date: '2020-04-17'
slug: am-i-less-happy-during-my-phd
categories: []
tags:
  - GCP
  - datavis
  - natural language processing
subtitle: ''
summary: ''
authors: []
lastmod: '2020-04-17T00:48:48+10:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---

# Introduction 

`tl;dr` The [googleLaguageR](https://github.com/ropensci/googleLanguageR) package is a powerful API to [Google's natural language processing](https://cloud.google.com/natural-language). One of the most intersting thing about this package is the sentiment analysis engine, which sends queries to Google's pre-trained sentiment analysis deep learning model. 

While doing a PhD can be a great experience for some, most PhD students have struggles and horror stories. I am more of the latter. 




```{r}
library(tidyverse)
library(lubridate)
```



```{r}
library(rvest)
library(xml2)
raw = xml2::read_html("~/Downloads/kevinwang_html/posts/your_posts_1.html")
raw_post = raw %>%
  html_nodes("div._2pi0:not(_3-95)") %>%
  html_text()


raw_time = raw %>% 
  html_nodes("div._3-94._2lem")
clean_time = raw_time %>% html_text()
```

```{r}
raw_tbl = tibble(post = raw_post, time = mdy_hm(clean_time)) %>% 
  dplyr::filter(year(time) > "2012") %>% 
  dplyr::mutate(
    contains_zh = post %>% str_detect("[^\x20-\x7F]"),
    zh_text = post %>% 
      str_replace_all("\n", " ") %>% 
      str_replace_all("[A-Za-z]|[:punct:]|\\s|\\d|[\u008f-\u009f]", ""),
    en_text = post %>% 
      str_replace_all("\n|[\u008f-\u009f]", " ") %>% 
      str_replace_all("Kevin Wang", " ") %>% 
      str_remove_all("[^\x20-\x7F]"))
```



```{r}
library(googleLanguageR)

english_gsent = googleLanguageR::gl_nlp(
  raw_tbl$en_text,
  nlp_type = "analyzeSentiment",
  language = "en")

chinese_gsent = googleLanguageR::gl_nlp(
  raw_tbl$zh_text,
  nlp_type = "analyzeSentiment",
  language = "zh")
```


<!-- ```{r} -->
<!-- save(english_gsent, chinese_gsent, file = "~/Desktop/sentiment.RData") -->
<!-- ``` -->

```{r}
load("~/Desktop/sentiment.RData")
```


```{r}
sent_tbl = raw_tbl %>% 
  dplyr::mutate(
    en_sent = english_gsent$documentSentiment$score,
    en_mag = english_gsent$documentSentiment$magnitude,
    zh_sent = chinese_gsent$documentSentiment$score,
    zh_mag = chinese_gsent$documentSentiment$magnitude,
    period = case_when(
      time <= "2015-03-01" ~ "Undergrad",
      time <= "2016-03-01" ~ "Honours",
      time <= "2019-09-01" ~ "PhD",
      TRUE ~ "Research Associate") %>% fct_relevel(c("Undergrad", "Honours", "PhD", "Research Associate")))
```


```{r}
theme_set(theme_classic(18))


sent_tbl %>% 
  dplyr::filter(contains_zh) %>% 
  dplyr::filter(!str_detect(en_text, "SUMS Executive")) %>% 
  ggplot(aes(x = en_sent, y = zh_sent)) +
  geom_point() +
  ggpubr::stat_cor() +
  geom_smooth(method = "lm") +
  facet_wrap(~period)
```


```{r}
library(patchwork)

sent_tbl %>% 
  ggplot() +
  geom_segment(aes(x = time, y = zh_sent, xend = time, yend = 0,
                   colour = period)
               , alpha = 0.5, size = 2) +
  geom_smooth(aes(x = time, y = zh_sent, colour = period)) +
  ylim(-1, 1) +
  scale_color_brewer(palette = "Set1") +
sent_tbl %>% 
  ggplot() +
  geom_segment(aes(x = time, y = en_sent, xend = time, yend = 0,
                   colour = period), 
               alpha = 0.5, size = 2) + 
  geom_smooth(aes(x = time, y = en_sent, colour = period)) +
  ylim(-1, 1) +
  scale_color_brewer(palette = "Set1") +
  patchwork::plot_layout(nrow = 2) 
```


```{r}
library(forecast)

ggCcf(x = sent_tbl$zh_sent, 
      y = sent_tbl$en_sent, lag.max = 10)
```





<!-- # Loading data from html -->
<!-- ```{r} -->
<!-- raw = fromJSON("~/Downloads/kevinwang_json_2/posts/your_posts_1.json", flatten = TRUE) %>% -->
<!--   janitor::clean_names() %>% -->
<!--   as_tibble() -->
<!-- ``` -->


<!-- ```{r} -->
<!-- text_only = raw %>% dplyr::select(timestamp, data) %>% -->
<!--   tidyr::unnest(data) %>% -->
<!--   dplyr::mutate( -->
<!--     timestamp = as.Date(as.POSIXct(timestamp, origin="1970-01-01"))) %>% -->
<!--   dplyr::filter(timestamp >= "2012-02-01", -->
<!--                 !str_detect(post, "birthday")) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- text_only$post[2] -->
<!-- ``` -->


<!-- ```{r} -->
<!-- text_only %>% -->
<!--   group_by(timestamp) %>% -->
<!--   tally() %>% -->
<!--   ggplot(aes(x = timestamp, y = n)) + -->
<!--   geom_path() -->
<!-- ``` -->


<!-- ```{r} -->
<!-- library(googleLanguageR) -->
<!-- g_raw = googleLanguageR::gl_nlp(text_only$post, -->
<!--                                 nlp_type = "analyzeSentiment", -->
<!--                                 language = "en") -->


<!-- g_sent = bind_cols(text_only, g_raw$documentSentiment) -->

<!-- g_sent %>% -->
<!--   dplyr::mutate(phd = timestamp >= "2016-03-01") %>% -->
<!--   ggplot(aes(x = timestamp, y = score)) + -->
<!--   geom_point(aes(size = magnitude)) + -->
<!--   geom_smooth(aes(group = phd, colour = phd), method = "loess") -->

<!-- g_sent %>% -->
<!--   dplyr::mutate( -->
<!--     phd = timestamp >= "2016-03-01") %>% -->
<!--   ggplot(aes(x = timestamp, y = magnitude)) + -->
<!--   geom_path() + -->
<!--   geom_smooth(aes(group = phd, colour = phd), method = "loess") -->
<!-- ``` -->


<!-- ```{r} -->
<!-- library(sentimentr) -->
<!-- library(furrr) -->
<!-- plan(multisession, workers = 5) -->

<!-- r_sent = text_only %>% -->
<!--   dplyr::mutate( -->
<!--     r_sent = furrr::future_map_dbl(post, ~ sentiment_by(.x)$ave_sentiment)) -->
<!-- ``` -->


<!-- ```{r} -->
<!-- r_sent %>% -->
<!--   dplyr::mutate( -->
<!--     phd = timestamp >= "2016-03-01") %>% -->
<!--   ggplot(aes(x = timestamp, y = r_sent)) + -->
<!--   geom_path() + -->
<!--   geom_smooth(aes(group = phd, colour = phd), method = "loess") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- set.seed(123) -->

<!-- r_sent_clust = r_sent %>% -->
<!--   dplyr::filter(timestamp >= "2016-03-01") %>% -->
<!--   dplyr::mutate(kmeans = kmeans(x = r_sent, centers = 2)$cluster) -->



<!-- library(patchwork) -->

<!-- r_sent %>% -->
<!--   dplyr::mutate( -->
<!--     period = case_when( -->
<!--       timestamp <= "2015-03-01" ~ "Undergrad", -->
<!--       timestamp <= "2016-03-01" ~ "Honours", -->
<!--       timestamp <= "2019-09-01" ~ "PhD", -->
<!--       TRUE ~ "Research Associate", -->
<!--       ) %>% fct_relevel(c("Undergrad", "Honours", "PhD", "Research Associate")) -->
<!--   ) %>% -->
<!--   group_by(period, year_month =  ymd(timestamp, truncated = 1)) %>% -->
<!--   tally() %>% -->
<!--   ggplot(aes(x = year_month, y = n, colour = period)) + -->
<!--   geom_point() + -->
<!--   theme(axis.text.x = element_text(angle = 90)) + -->
<!--   labs(x = "Time", -->
<!--        y = "", -->
<!--        title = "Number of posts per month") + -->
<!--   scale_color_brewer(palette = "Set1") + -->
<!--   theme_classic(18) + -->
<!--   r_sent %>% -->
<!--   dplyr::mutate( -->
<!--     period = case_when( -->
<!--       timestamp <= "2015-03-01" ~ "Undergrad", -->
<!--       timestamp <= "2016-03-01" ~ "Honours", -->
<!--       timestamp <= "2019-09-01" ~ "PhD", -->
<!--       TRUE ~ "Research Associate", -->
<!--       ) %>% fct_relevel(c("Undergrad", "Honours", "PhD", "Research Associate")), -->
<!--     post_length = purrr::map_int(post, str_length) -->
<!--   ) %>% -->
<!--   group_by(period, year_month =  ymd(timestamp, truncated = 1)) %>% -->
<!--   summarise(mean_post_length = mean(post_length)) %>% -->
<!--   ggplot(aes(x = year_month, y = mean_post_length, colour = period)) + -->
<!--   geom_point() + -->
<!--   theme(axis.text.x = element_text(angle = 90)) + -->
<!--   labs(x = "Time", -->
<!--        y = "", -->
<!--        title = "Avg. length of post") + -->
<!--   scale_color_brewer(palette = "Set1") + -->
<!--   theme_classic(18) + -->
<!--   r_sent %>% -->
<!--   dplyr::mutate( -->
<!--     period = case_when( -->
<!--       timestamp <= "2015-03-01" ~ "Undergrad", -->
<!--       timestamp <= "2016-03-01" ~ "Honours", -->
<!--       timestamp <= "2019-09-01" ~ "PhD", -->
<!--       TRUE ~ "Research Associate", -->
<!--       ) %>% fct_relevel(c("Undergrad", "Honours", "PhD", "Research Associate")) -->
<!--   ) %>% -->
<!--   ggplot(aes(x = timestamp, y = r_sent)) + -->
<!--   geom_path() + -->
<!--   geom_smooth(aes(group = period, colour = period), method = "loess") + -->
<!--   labs(x = "Time", -->
<!--        y = "", -->
<!--        title = "`sentimentr` sentiment scoring") + -->
<!--   scale_color_brewer(palette = "Set1") + -->
<!--   theme_classic(18) + -->
<!--   patchwork::plot_layout(nrow = 3) -->
<!--   # geom_rug(data = r_sent_clust, aes(colour = factor(kmeans))) -->
<!-- ``` -->


