---
title: "NASA - FIRMS fire spread around Nowra"
author: "Kevin Wang"
date: "3 Jan 2020"
output: 
  html_document:
    code_download: true
    toc: true
    theme: paper
---

# Summary

This is a document for visualising the bushfire spread around my family home in Nowra, NSW, Australia. 

The data was obtained from [NASA - FIRMS Archive Downloads](https://firms.modaps.eosdis.nasa.gov/download/).The frozen copy of this data (accessed on the 3rd of Jan 2020, restricted to Australia, range from 1st Dec to 3rd of Jan 2020) that I obtained can be found [here](https://kevinwang09.github.io/tidytuesday/bushfire/DL_FIRE_M6_94441/fire_nrt_M6_94441.csv).

Parts of this code will not be reproducible due to the `ggmap` Google API. You might want to consult the manual of this package to obtain the instructions of setting up your laptop with the API.  


# Loading 

## Packages
```{r, massage = FALSE}
library(tidyverse)
library(ggmap)
library(lubridate)
library(gganimate)
```


## Loading data

```{r}
raw = readr::read_csv("DL_FIRE_M6_94441/fire_nrt_M6_94441.csv")

glimpse(raw)

raw$acq_time %>% unique %>% sample(size = 50) ## Four digits time stamp
```

# Basic data cleaning and variable selection

I am only selecting some variables of interest to me. 

```{r}
clean = raw %>% 
  dplyr::transmute(
    latitude,
    longitude,
    brightness,
    confidence,
    bright_t31,
    frp,
    daynight,
    my_time = paste0(acq_date, " ", acq_time) %>% 
      lubridate::ymd_hm(), ## Aftering pasting the hours and minites, we use lubridate to clean this 
    confidence,
    conf_cat = cut(
      confidence, 
      c(-1, 50, 101), 
      labels = c("low", "high")) %>% as.factor,
    brightness,
    bright_cat = cut(
      brightness, 
      c(300, 350, 400, Inf), 
      labels = c("low", "med", "high")) %>% as.factor
  )

glimpse(clean)
```


# Google Maps

This code is not reproducible unless you have the correct Google Maps API. 

```{r}
nowra_map <- get_map("Nowra Hill, NSW", zoom = 10)
str(nowra_map)
map_bounds = attr(nowra_map, "bb")
```


## Subsetting the fire data to only those in the map
```{r}
nowra_data = clean %>% 
  dplyr::filter(
    latitude >= map_bounds$ll.lat,
    latitude <= map_bounds$ur.lat,
    longitude >= map_bounds$ll.lon,
    longitude <= map_bounds$ur.lon)

nowra_data %>% glimpse
```


```{r}
p1 = ggmap(nowra_map) +
  geom_point(data = nowra_data, 
             aes(x = longitude,
                 y = latitude,
                 colour = brightness)) +
  geom_point(x = 150.5934431,
             y = -34.8432388,
             colour = "blue", size = 2) + ## A location near my home
  scale_colour_distiller(palette = "Reds", direction = 1) +
  transition_time(my_time) +
  labs(title = "Time: {frame_time}") +
    shadow_trail(distance = 0.05, colour = "black", alpha = 1)

p2 = animate(p1, nframes = 100)

p2
```


```{r, echo = FALSE}
anim_save(filename = "./Nowra_fire.gif", 
          animation = p2)
```



```{r}
sessionInfo()
```

