---
title: "Comparing different "
author: "Kevin Wang"
date: "07 Dec 2019"
output: 
  html_document:
    code_download: true
    toc: true
    theme: paper
---


# Aim of this document

This is a demonstration of the `ggiraph` package's interactive visualisation capability. Also, since `patchwork` was recently put on CRAN, "side-by-side interactive plotting" just got easier!

I will perform the PCA, tSNE and UMAP dimensionality reduction on the `iris` data. 

Just so you know that I know:

+ Yes, `iris` data is overused, but hey! It is only a quick coding demo!

+ Yes, I know about `plotly`, it also has the side-by-side option, but it is very painful when there are multiple plotting aesthetics.

```{r, message=FALSE}
library(ggiraph)
library(patchwork)
library(tidyverse)
library(Rtsne)
library(umap)
```

# Computing the dimensionality reduction objects

```{r}
## We:
## + convert the iris data to a tibble
## + remove one duplicated row
## + add unique identifier to each row

iris_tbl = iris %>% 
  as_tibble() %>% 
  distinct() %>% 
  tibble::rownames_to_column("id")

iris_tbl

## Computing the PCA, TSNE and UMAP objects
iris_pca = prcomp(iris_tbl %>% select_if(is.numeric))
iris_tsne = Rtsne(iris_tbl %>% select_if(is.numeric))
iris_umap = umap(iris_tbl %>% select_if(is.numeric))
```

# Preparing the plotting data and static plots 

The main function we will use is `geom_point_interactive` from the `ggiraph` package. It behaves just like `geom_point`, but with two additional aesthetics. The `data_id` aesthetic is critical to link observations between plots and the `tooltip` aesthetic is optional but nice to have when mouse over a point. After making these plots, the `girafe` function using the same syntax in `patchwork` will allow us to make a pretty interactive plot!

```{r}
## Getting two reduced dimensions for each method
iris_plotdf = iris_tbl %>% 
  mutate(
    pca_1 = iris_pca$x[,1],
    pca_2 = iris_pca$x[,2],
    tsne_1 = iris_tsne$Y[,1],
    tsne_2 = iris_tsne$Y[,2],
    umap_1 = iris_umap$layout[,1],
    umap_2 = iris_umap$layout[,2]
  )



p1 = iris_plotdf %>% 
  ggplot(aes(x = pca_1, y = pca_2, 
             colour = Species,
             tooltip = id,
             data_id = id)) +
  geom_point_interactive() +
  labs(title = "PCA")

p2 = iris_plotdf %>% 
  ggplot(aes(x = tsne_1, y = tsne_2, 
             colour = Species,
             tooltip = id,
             data_id = id)) +
  geom_point_interactive() +
  labs(title = "tSNE")

p3 = iris_plotdf %>% 
  ggplot(aes(x = umap_1, y = umap_2, 
             colour = Species,
             tooltip = id,
             data_id = id)) +
  geom_point_interactive() +
  labs(title = "UMAP")
```



```{r, fig.width = 15, fig.height=5}
p1 + p2 + p3
```


At this point, `p1 + p2 + p3` are combined together as a static plot thanks to `patchwork`. But making it iterative is just one line. 

## The magical interactive visualisation
```{r}
girafe(code = print(p1 + p2 + p3), 
       width_svg = 15, height_svg = 5)
```



## Extensions

For the `tidyverse` fanatics (myself included), you will see that we have made three plots with similar structures and cringe a little bit. Here is another way to make similar plot by pivoting. 

```{r, fig.width = 15, fig.height=5}
iris_long_plotdf = iris_plotdf %>% 
  tidyr::pivot_longer(cols = matches("1|2"), 
                      names_to = c("method", "dim"),
                      names_sep = "_") %>% 
  tidyr::pivot_wider(names_from = dim, 
                     values_from = value)

iris_long_plotdf

p4 = iris_long_plotdf %>% 
  ggplot(aes(x = `1`, y = `2`, 
             colour = Species,
             tooltip = id,
             data_id = id)) +
  geom_point_interactive() +
  facet_grid(~method, scales = "free")

girafe(code = print(p4),
       width_svg = 10, height_svg = 5)
```



# Session Info
```{r}
sessionInfo()
```

