---
title: "side-by-side (gg)plots in R"
author: "Kevin Wang"
date: "07 Dec 2019"
output: 
  html_document:
    code_download: true
    toc: true
    theme: paper
---

# Aim of this document

I will show you some options to make side-by-side ggplots in R. While there are many options t=out there, I will cover only three: 

+ `patchwork`, which allows combination of two or more ggplots using a natural syntax of `+` (similar output as `gridExtra::grid.arrange`)
+ `ggiraph`, which leverages `patchwork`'s syntax to make **interactive** ggplots!
+ `plotly`, which converts ggplots into interactive plotly-plots before combining everything into subplots. 

See more [here](dim_redu.html)!


## A small note

I wish to demonstrate a rather annoying behaviour of `ggplotly` in the `plotly` package. 

One of the most powerful function in `plotly` is the `ggplotly` function. For a `ggplot2`-addict like myself it takes a static ggplot and makes it interative. 

However, this function has an undesirable side-effect when trying to produce a side-by-side plot with mouse-over functionality. I found that the work around of this to be a bit harder than I would like. One way around this is to use the `ggiraph` package. I will demonstrate both of these packages below. 


Just so you know that I know:
+ Yes, `mtcars` data is overused, but hey! It is only a quick coding demo!

```{r, message=FALSE}
library(tidyverse)
library(ggiraph)
library(patchwork)
library(plotly)
```

# Making a better `mtcars` data

The code below will make variables with less than 6 unique values to be factor variables. This is because variables like `cyl` should be visualised as factors. 

```{r}
## We will make the mtcars data into a tibble and convert the rownames into a column of id
dat = mtcars %>% 
  tibble::rownames_to_column("id") %>% 
  as_tibble()

## Computing the number of unique values in each column
## We will make columns with less or equal to 6 unique values to be factors
num_unique = function(x){x %>% unique %>% length}
(vars_unique = dat %>% apply(2, num_unique))
(fct_cols = names(vars_unique)[vars_unique <= 6])

## Mutating the columns based on the previous calculation
dat = dat %>% 
  as_tibble() %>% 
  mutate_at(.vars = fct_cols,
            .funs = as.factor)

dat
```


# Making individual static plots

These are standard ggplots.

```{r, fig.width = 6, fig.height=4}
g1 = dat %>% 
  ggplot(aes(x = mpg, y = disp, 
             colour = cyl,
             shape = vs,
             data_id = id,
             tooltip = id)) +
  geom_point_interactive(size = 4) +
  labs(x = "Miles per gallon", 
       y = "Displacement")

g1
```

```{r}
g2 = dat %>% 
  ggplot(aes(x = mpg, y = drat, 
             colour = cyl,
             shape = vs, 
             data_id = id,
             tooltip = id)) +
  geom_point_interactive(size = 4) +
  labs(x = "Miles per gallon", 
       y = "Rear axle ratio")
g2
```



# Side-by-side static plots using `patchwork`

After some wait, `patchwork` is finally on CRAN! It's great advantage is putting together `p1` and `p2` using the familiar `ggplot2` syntax of `+`. 



<blockquote class="twitter-tweet"><p lang="en" dir="ltr">So so excited 🎉🎉🎉 patchwork is now on CRAN <a href="https://t.co/3kieCQxoJb">https://t.co/3kieCQxoJb</a></p>&mdash; Thomas Lin Pedersen (@thomasp85) <a href="https://twitter.com/thomasp85/status/1201125936311275522?ref_src=twsrc%5Etfw">December 1, 2019</a></blockquote> <script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>

```{r, fig.width = 12, fig.height=4}
g1 + g2
```


Note: even though we designed these two plots as interactive plots, they will not become interactive until we put a small wrapper around it, which we will see later. 


# Side-by-side interactive plots using `ggiraph`

The main function we will use is `geom_point_interactive` from the `ggiraph` package. It behaves just like `geom_point`, but with two additional aesthetics. The `data_id` aesthetic is critical to link observations between plots and the `tooltip` aesthetic is optional but nice to have when mouse over a point. After making these plots, the `girafe` function using the same syntax in `patchwork` will allow us to make a pretty interactive plot!


Since I am writing in a RMarkdown file, I will also specify the height and width of my output using optional parameters.

```{r}
girafe(code = print(g1 + g2), width_svg = 12, height_svg = 4)
```


# Side-by-side interactive plots using `plotly`

In the code below, `g3` and `g4` are identical to `g1` and `g2` respectively, except that `geom_point` was used. The `ggplotly` function will convert the two plots into plotly interactive plots and the `subplot` will combine the two plots into one singular plot.


```{r}
g3 = dat %>%
  ggplot(aes(x = mpg, y = disp,
             colour = cyl,
             shape = vs,
             data_id = id,
             tooltip = id)) +
  geom_point(size = 4) +
  labs(x = "Miles per gallon",
       y = "Displacement")


g4 = dat %>%
  ggplot(aes(x = mpg, y = drat,
             colour = cyl,
             shape = vs,
             data_id = id,
             tooltip = id)) +
  geom_point(size = 4) +
  labs(x = "Miles per gallon",
       y = "Rear axle ratio")

subplot(ggplotly(g3),
        ggplotly(g4))
```


Look at the colour and shape legend being repeated. This is the main reason that I don't prefer `ggplotly` for this type of task. But don't get me wrong! I still use `ggplotly` in my everyday work, but just not this task.


# Session info
```{r}
sessionInfo()
```

